### 2017-03-26 paper/oopsla17-submission/gallery.tex
failed predictions

\paragraph{Extracting the Digits of an Integer}
Our first program is a simple recursive function |digitsOfInt| that
extracts the digits of an |int|.
%
\begin{ecode}
  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      [n mod 10] @ __[ ___=digitsOfInt (n / 10)=___ ]__
\end{ecode}
%
Unfortunately, the student has decided to wrap the recursive call to
|digitsOfInt| with a list literal, even though |digitsOfInt| already
returns an |int list|.
%
Thus, the list literal is inferred to have type |int list list|, which
is incompatible with the |int list| on the left of the |@| (list append)
operator.
%
Both \sherrloc and the \ocaml compiler blame the recursive call for
returning a |int list| rather than |int|, but the recursive call is
actually correct!
%
As our Decision Tree correctly points out, the fault lies with the list
literal \emph{surrounding} the recursive call, remove it and the type
error disappears.
%
\ES{TODO: WHY?}

\paragraph{Padding a list}
Our next program, |padZero|, is given two |int list|s as input, and must
left-pad the shorter one with enough zeros that the two output lists
have equal length.

@@

\paragraph{Computing the Fixed Point of a Function}
Finally, our students must write a |fixpoint| function that computes the
fixed point of a given function |f|, starting from a base case |b|.
%
As a hint we first have them write a |wwhile| function that performs the
functional equivalent of a while-loop, repeatedly passing a function's
output back in until it receives a (boolean) signal to stop.
%
\begin{ecode}
  let rec wwhile (f, b) =
    match f b with
    | (x, false) -> x
    | (x, true)  -> wwhile (f, x)
\end{ecode}
\lstset{firstnumber=last}
%
The |fixpoint| function can then be written as a clever instantiation of
the arguments to |wwhile|.
%
\begin{ecode}
  let fixpoint (f, b) =
    let g = __let bb = f b in ___=(bb, (bb = b))=_ in
    wwhile (g, b)
\end{ecode}
\lstset{firstnumber=1}
%
Sadly, our student has forgotten that |g| should itself be a function.
%
As a result, she passes a pair of a \emph{pair} and a starting value to
|wwhile|, rather than a pair of a \emph{function} and a starting value.
%
\sherrloc deduces that the call to |wwhile| is likely correct (\ocaml
actually blames the use of |g|), but identifies the construction of the
pair inside |g| as the most likely culprit, while our Decision Tree
correctly identifies the \emph{definition} of |g| as the source of the
error.
\ES{TODO: WHY?}

>>>>

\paragraph{Extracting the Digits of an Integer}
% data/sp14/1655.ml
Our first program is a simple recursive function |digitsOfInt| that
extracts the digits of an |int|.
%
\begin{ecode}
  let rec digitsOfInt n =
    if n <= 0 then
      []
    else
      [n mod 10] @ __[ ___=digitsOfInt (n / 10)=___ ]__
\end{ecode}
%
Unfortunately, the student has decided to wrap the recursive call to
|digitsOfInt| with a list literal, even though |digitsOfInt| already
returns an |int list|.
%
Thus, the list literal is inferred to have type |int list list|, which
is incompatible with the |int list| on the left of the |@| (list append)
operator.
%
Both \sherrloc and the \ocaml compiler blame the recursive call for
returning a |int list| rather than |int|, but the recursive call is
actually correct!
%
As our Decision Tree correctly points out, the fault lies with the list
literal \emph{surrounding} the recursive call, remove it and the type
error disappears.
%
\ES{TODO: WHY?}

\paragraph{Padding a list}
% data/sp14/0306.ml
Our next program, |padZero|, is given two |int list|s as input, and must
left-pad the shorter one with enough zeros that the two output lists
have equal length.

@@

\paragraph{Computing the Fixed Point of a Function}
% data/sp14/0941.ml
Finally, our students must write a |fixpoint| function that computes the
fixed point of a given function |f|, starting from an initial value |b|.
%
As a hint we first have them write a |wwhile| function that performs the
functional equivalent of a while-loop, repeatedly passing a function's
output back in until it receives a (boolean) signal to stop.
%
\begin{ecode}
  let rec wwhile (f, b) =
    match f b with
    | (x, false) -> x
    | (x, true)  -> wwhile (f, x)
\end{ecode}
\lstset{firstnumber=last}
%
The |fixpoint| function can then be written as a clever instantiation of
the arguments to |wwhile|.
%
\begin{ecode}
  let fixpoint (f, b) =
    let g = __let bb = f b in ___=(bb, (bb = b))=_ in
    wwhile (g, b)
\end{ecode}
\lstset{firstnumber=1}
%
Sadly, our student has forgotten that |g| should itself be a function.
%
As a result, she passes a pair of a \emph{pair} and a starting value to
|wwhile|, rather than a pair of a \emph{function} and a starting value.
%
\sherrloc deduces that the call to |wwhile| is likely correct (\ocaml
actually blames the use of |g|), but identifies the construction of the
pair inside |g| as the most likely culprit, while our Decision Tree
correctly identifies the \emph{definition} of |g| as the source of the
error.
\ES{TODO: WHY?}

\subsubsection{Failed Predictions}
\label{sec:failed-predictions}
Of course, our classifiers are sometimes wrong, we focus next on the
10\% of programs where \emph{none} of our Top-3 predictions are correct.

% data/sp14/0148.ml
\begin{ecode}
  let rec clone x n =
    let loop acc n =
      if n <= 0 then
        acc
      else
        clone ==([==_=x=_==] @ acc)== (n - 1) in
    loop [] n
\end{ecode}
\ES{TODO: our 2nd prediction matches \sherrloc and \ocaml (occurs check), correct fix is to replace recursive call to clone with loop}

% data/sp14/0234.ml
\begin{ecode}
  let pipe fs =
    let f a x z = _=z=___ a__ in
    let base y = y in
    List.fold_left f base fs
\end{ecode}
\ES{TODO: fix is to swap a and z, \sherrloc is correct, \ocaml blames use of f in fold-left..}

% data/sp14/0255.ml
\begin{ecode}
  let rec wwhile (f, b) =
    let check = f b in
    let (x, y) = (f, b) in
    if __y__ = false then
      x
    else
      wwhile (f, ==x==)
\end{ecode}
\ES{marking ``y'' seems totally bogus, why is it even part of the error
  slice? (other locations are ``f b'' and ``false'' though, whihc is not
  any better)}
